---
title: "Homework 2"
author: "Anusorn Thanataveerat"
date: '2018-09-25'
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

# Problem 1

Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable

```{r Load_Package, include = FALSE}
library(tidyverse)
library(gridExtra)
library(janitor)
```
```{r problem_1, echo = TRUE}
original_dat <-
  as.tibble(
  read.csv(
  'http://p8105.com/data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv'
  )
  ) %>% clean_names()
  
  subset_dat <-
  original_dat %>% select(line:entry, vending, ada, ada_notes) %>%
  mutate(entry = ifelse(entry == 'YES', TRUE, FALSE))
  dedup_subset_dat <-
  subset_dat %>% distinct(line, station_name, .keep_all = TRUE)
```

Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

**Answer**:Once downloaded, the data set consists of `r ncol(original_dat)` variables and `r nrow(original_dat)` observations. We further included only the variables as suggested and the number of variables came down to `r ncol(subset_dat)`. At this point, we noticed rows with identical input of a combination of *line* and *station name* thus we de-duplicated based on those two variables and ended up with **`r nrow(dedup_subset_dat)`** unique stations in the NYC subway system. There are `r sum(dedup_subset_dat$ada)` stations (`r round(100* sum(dedup_subset_dat$ada)/nrow(dedup_subset_dat),2)` %) which are ADA compliant. Also,
What proportion of station entrances / exits without vending allow entrance?




**Answer**: The character and factor variables can be converted to numeric variables.
Again, it converts `TRUE` to 1 and `FALSE` to 0 for logic variable. For character variable, the numeric values take values from its 'face' value. For factor variable, the results are the levels of the factor. For example in this case there are 7 levels and the with number *500* is treated as the highest level, therefore it is recoded as 7 while *7* is the fifth highest level and is recoded as 5.

In a second code chunk, convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?




